Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi

### 2013-10-04 writeup/evaluation.tex
More edits'

\vspace{.5em}\noindent\textbf{Non-serializability} Most existing
distributed implementations of weak isolation are not highly
available. Lock-based mechanisms such as those in Gray's original
proposal~\cite{gray-isolation} do not degrade gracefully in the
presence of partial failures. (Note, however, that lock-based
protocols \textit{do} offer the benefit of recency guarantees.) While
multi-versioned storage systems allow for a variety of transactional
guarantees, few offer traditional weak isolation (e.g.,
non-``tentative update'' schemes) in this context.  Chan and Gray's
read-only transactions have item-cut isolation with causal consistency
and MAV (session \textit{PL-2L}~\cite{adya}) but are unavailable in
the presence of coordinator failure and assume serializable update
transactions~\cite{readonly}; this is similar to read-only and
write-only transactions more recently proposed by Eiger~\cite{eiger}.
Brantner's S3 database~\cite{kraska-s3} and
Bayou~\cite{sessionguarantees} can all provide variants of session
\textit{PL-2L} with high availability, but none provide this HAT
functionality without substantial modification. Swift~\cite{swift} and
bolt-on causal consistency~\cite{bolton} are closest to providing the
maximal sticky HAT semantics we have considered. Accordingly, it is
possible to implement many guarantees weaker than
serializability---including guarantees that are achievable with high
availability---and still not achieve high availability. We view high
availability as a core design consideration in future concurrency
control designs.

>>>>

\vspace{.5em}\noindent\textbf{Non-serializability} Most existing
distributed implementations of weak isolation are not highly
available. Lock-based mechanisms such as those in Gray's original
proposal~\cite{gray-isolation} do not degrade gracefully in the
presence of partial failures. (Note, however, that lock-based
protocols \textit{do} offer the benefit of recency guarantees.) While
multi-versioned storage systems allow for a variety of transactional
guarantees, few offer traditional weak isolation (e.g.,
non-``tentative update'' schemes) in this context.  Chan and Gray's
read-only transactions have item-cut isolation with causal consistency
and MAV (session \textit{PL-2L}~\cite{adya}) but are unavailable in
the presence of coordinator failure and assume serializable update
transactions~\cite{readonly}; this is similar to read-only and
write-only transactions more recently proposed by Eiger~\cite{eiger}.
Brantner's S3 database~\cite{kraska-s3} and
Bayou~\cite{sessionguarantees} can all provide variants of session
\textit{PL-2L} with high availability, but none provide this HAT
functionality without substantial modification. Accordingly, it is
possible to implement many guarantees weaker than
serializability---including HAT semantics---and still not achieve high
availability. We view high availability as a core design consideration
in future concurrency control designs.

### 2013-10-04 writeup/hat-analysis.tex
More edits'

As a straw-man, replicas can store all versions ever written to each
data item. Replicas can gossip information about versions they have
observed and construct a lower bound on the versions that can be found
on each replica. At the start of each transaction, clients choose a
bound that is guaranteed to be lower than or equal to the this global
lower bound, and, during transaction execution, replicas return the
latest version of each item that is not greater than the client's
chosen bound. If the lower bound is advanced along transactional
boundaries, clients will observe MAV. This algorithm has several
variants in the literature~\cite{readonly, swift}: bounds can be
represented by vector clocks, and older versions can be asynchronously
garbage collected.

@@

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)
\\T_2 &: r_x(a)\vspace{-1em}
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, then, for transactional availability,
the server must allow $T_1$ to commit. If the same client subsequently
executes $T_2$ against the same (partitioned) server in the same
session, then it will be able to read its writes. However, if the
network topology changes and the client can only execute $T_2$ on a
different replica that is partitioned from the replica that executed
$T_1$, then the system will have to either stall indefinitely to allow
the client to read her writes (violating transactional availability)
or will have to sacrifice read your writes guarantees. However, if the
client remains sticky with the server that executed $T_1$, then we can
disallow this scenario. Accordingly, read your writes, and, by proxy,
causal consistency and PRAM require stickiness. Read your writes is
provided by default in a sticky system. Causality and PRAM guarantees
can be accomplished with well-known variants~\cite{causalmemory,
  bolton, eiger, sessionguarantees, swift} of the prior session
guarantee algorithms we presented earlier: only reveal new writes to
clients when their dependencies have been revealed.

@@

\vspace{.5em}\noindent{\textbf{Convergence}} To require that the
system propagates writes between replicas, we can require convergence,
or \textit{eventual consistency}: in the absence of new mutations to a
data item, in the absence of partitions, all servers should eventually
agree on the value for each item~\cite{cac, vogels-defs}. This is
typically accomplished by any number of anti-entropy protocols, which
periodically update neighboring servers with the latest value for each
data item~\cite{antientropy}. Establishing a final convergent value is
related to determining a total order on transaction updates to each
item, as in Read Uncommitted.

>>>>

As a straw-man, replicas can store all versions ever written to each
data item. Replicas can gossip information about versions they have
observed and construct a lower bound on the versions that can be found
on every replica (which can be represented by either a list of
versions, or, more realistically, a vector clock). At the start of
each transaction, clients can choose a \textit{read timestamp} that is
lower than or equal to the this global lower bound, and, during
transaction execution, replicas return the latest version of each item
that is not greater than the client's chosen timestamp. If this lower
bound is advanced along transactional boundaries, clients will observe
MAV. This algorithm has several variants in the
literature~\cite{readonly, swift}: bounds can be represented by vector
clocks, and older versions can be asynchronously garbage collected.

@@

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions:
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: r_x(a)
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, then, for transactional availability,
the server must allow $T_1$ to commit. If the same client subsequently
executes $T_2$ against the same (partitioned) server in the same
session, then it will be able to read its writes. However, if the
network topology changes and the client can only execute $T_2$ on a
different replica that is partitioned from the replica that executed
$T_1$, then the system will have to either stall indefinitely to allow
the client to read her writes (violating transactional availability)
or will have to sacrifice read your writes guarantees. However, if the
client remains sticky with the server that executed $T_1$, then we can
disallow this scenario. Accordingly, read your writes, and, by proxy,
causal consistency and PRAM require stickiness. Read your writes is
provided by default in a sticky system. Causality and PRAM guarantees
can be accomplished with well-known variants~\cite{causalmemory,
  bolton, eiger, sessionguarantees, swift} of the prior session
guarantee algorithms we presented earlier: only reveal new writes to
clients when their dependencies have been revealed.

@@

\vspace{.5em}\noindent{\textbf{Convergence} Under arbitrary (but not
  infinite delays), HAT systems can ensure convergence, or
  \textit{eventual consistency}: in the absence of new mutations to a
  data item, all servers should eventually agree on the value for each
  item~\cite{cac, vogels-defs}. This is typically accomplished by any
  number of anti-entropy protocols, which periodically update
  neighboring servers with the latest value for each data
  item~\cite{antientropy}. Establishing a final convergent value is
  related to determining a total order on transaction updates to each
  item, as in Read Uncommitted.

### 2013-10-04 writeup/hat-vldb.tex
More edits'

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
\usepackage{draftwatermarkbottom, draftwatermarktop}
\usepackage{scrextend, MnSymbol}
\usepackage[hyphens]{url}

@@

@@

>>>>

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
%\usepackage{draftwatermarkbottom, draftwatermarktop}
\usepackage{scrextend, MnSymbol}
\usepackage[hyphens]{url}

@@

\setlength{\abovedisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{4pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\belowdisplayshortskip}{4pt}

@@

\section*{Acknowledgments}
We would like to thank Peter Alvaro, Neil Conway, Evan Jones, Adam
Oliner, Aurojit Panda, Shivaram Venkataraman, and the HotOS and VLDB
reviewers for their helpful feedback on earlier revisions of this
work. This research was supported in part by the Air Force Office of
Scientiﬁc Research (grant FA95500810352), DARPA XData Award
FA8750-12-2-0331, the National Science Foundation (grants CNS-0722077,
IIS-0713661, and IIS-0803690), NSF CISE Expeditions award CCF-1139158,
the National Science Foundation Graduate Research Fellowship (grant
DGE-1106400), and by gifts from Amazon Web Services, Google, SAP,
Cisco, Clearstory Data, Cloudera, Ericsson, Facebook, FitWave, General
Electric, Hortonworks, Huawei, Intel, Microsoft, NetApp, Oracle,
Samsung, Splunk, VMware, WANdisco, and Yahoo!.

### 2013-10-04 writeup/intro.tex
More edits'

Our investigation is based on both impossibility results and several
constructive, proof-of-concept algorithms. For example, Snapshot
Isolation and Repeatable Read isolation are not HAT-compliant because
they require detecting conflicts between concurrent updates (as needed
for preventing Lost Updates or Write Skew phenomena), which we show is
unavailable. However, Read Committed isolation, transactional
atomicity (Read Atomic isolation), and many of the weak consistency
models from database and distributed systems are achievable via
algorithms that rely on multi-versioning and limited client-side
caching. For several guarantees, such as causal consistency with
phantom prevention and ANSI Repeatable Read, we consider a modified
form of high availability in which clients ``stick to'' (i.e., have
affinity with) at least one server---a property which is often
implicit in the distributed systems literature~\cite{herlihy-art,
  eiger, cac} but which requires explicit consideration in a
client-server replicated database context. This sticky availability is
widely employed~\cite{eiger, vogels-defs} but is a less restrictive
model (and therefore more easily achievable) than traditional high
availability.

@@

We recognize that the large variety of ACID isolation levels and
distributed consistency models (and therefore in our taxonomization)
can be confusing and that the subtle distinctions between them may
appear to be primarily of academic concern. Accordingly, we offer the
following pragmatic takeaways:
\begin{introenumerate}
\item The default (and sometimes strongest) configurations of most
  widely-deployed database systems expose a range of anomalies that
  can compromise application-level consistency.

@@

\item In addition to providing a guaranteed response and horizontal
  scale-out, these HAT models allow for one to three order of
  magnitude lower latencies on current infrastructure.

>>>>

Our investigation is based on both impossibility results and several
constructive, proof-of-concept algorithms. For example, Snapshot
Isolation and Repeatable Read isolation are not HAT-compliant because
they require detecting conflicts between concurrent updates (as needed
for preventing Lost Updates or Write Skew phenomena), which we show is
unavailable. However, Read Committed isolation, transactional
atomicity (``Monotonic Atomic View'' isolation), and many other consistency
models from database and distributed systems are achievable via
algorithms that rely on multi-versioning and limited client-side
caching. For several guarantees, such as causal consistency with
phantom prevention and ANSI Repeatable Read, we consider a modified
form of high availability in which clients ``stick to'' (i.e., have
affinity with) at least one server---a property which is often
implicit in the distributed systems literature~\cite{herlihy-art,
  eiger, cac} but which requires explicit consideration in a
client-server replicated database context. This sticky availability is
widely employed~\cite{eiger, vogels-defs} but is a less restrictive
model (and therefore more easily achievable) than traditional high
availability.

@@

We recognize that the large variety of ACID isolation levels and
distributed consistency models (and therefore in our taxonomization)
can be confusing and that the subtle distinctions between them can
often appear to be primarily of academic concern. Accordingly, we
offer the following pragmatic takeaways:
\begin{introenumerate}
\item The default (and sometimes strongest) configurations of most
  widely-deployed database systems expose a range of anomalies that
  can compromise application-level consistency.

@@

\item In addition to providing a guaranteed response and horizontal
  scale-out, these highly available HAT models allow one to three
  order of magnitude lower latencies on current infrastructure.

### 2013-10-04 writeup/relatedwork.tex
More edits'

Weak consistency and high availability have been well
studied. Serializability has long been known to be
unachievable~\cite{davidson-survey} and Brewer's CAP Theorem has
attracted considerable attention~\cite{gilbert-cap}. Recent work on
PACELC expands CAP by considering connections between ``weak
consistency'' and low latency~\cite{abadi-pacelc}, while several
studies examine weak isolation guarantees~\cite{adya,
  ansicritique}. There are a wide range of coordination-avoiding
``optimistic replication'' strategies~\cite{optimistic} and several
recent efforts at further taxonomizing these strategies in light of
current practice and the proliferation of ``eventually consistent''
stores~\cite{bailis-ec, bernstein-survey} (notably, Bernstein and
Das~\cite{bernstein-survey} specifically highlight the importance of
stickiness~\cite{sessionguarantees, vogels-defs}---which we formalize
in Section~\ref{sec:sticky}).  However, as we have discussed, the
connection between weak, multi-object consistency models in a
distributed setting and high availability is relatively
unexplored. With the exception of our preliminary workshop paper
discussing transactional availability, real-world ACID, and HAT RC and
I-CI~\cite{hat-hotos}---which this work expands, providing additional
algorithms and impossibility results, a full taxonomization,
experimental and application analysis, and a greater discussion of
implications and related work---we believe this paper is the first to
explore this area.

>>>>

Weak consistency and high availability have been well
studied. Serializability has long been known to be
unachievable~\cite{davidson-survey} and Brewer's CAP Theorem has
attracted considerable attention~\cite{gilbert-cap}. Recent work on
PACELC expands CAP by considering connections between ``weak
consistency'' and low latency~\cite{abadi-pacelc}, while several
studies examine weak isolation guarantees~\cite{adya,
  ansicritique}. There are a wide range of coordination-avoiding
``optimistic replication'' strategies~\cite{optimistic} and several
recent efforts at further taxonomizing these strategies in light of
current practice and the proliferation of ``eventually consistent''
stores~\cite{bailis-ec, bernstein-survey} (notably, Bernstein and
Das~\cite{bernstein-survey} specifically highlight the importance of
stickiness~\cite{sessionguarantees, vogels-defs}---which we formalize
in Section~\ref{sec:sticky}).  Aside from our earlier workshop paper
discussing transactional availability, real-world ACID, and HAT RC and
I-CI~\cite{hat-hotos}---which this work expands with additional
semantics, algorithms, and analysis---we believe this paper is the
first to explore the connections between transactional semantics, data
consistency, and (Gilbert and Lynch~\cite{gilbert-cap}) availability.
